---
title: "CAMDA2024 - Microbiom"
subtitle: "002_Gut"
author: "David Alberto García Estrada"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We adjust the working area

```{r}
# Set seed
set.seed(1992)

# Display current directory
getwd()
# If necessary, change the path where the stored files are located. 
workingDir <- "."
setwd(workingDir)

# Type data
data <- "Gut"

# Directories
D.dir <- paste0(Data, data, "/")
R.dir <- paste0(Results, data, "/")
```

# Data

Import data

```{r}
# List files
list.files(D.dir)

# Metadata
Meta <- read.table(paste0(D.dir, "metadata.txt"), header = TRUE)
rownames(Meta) <- Meta$SampleID
Meta$Diagnosis <- as.factor(Meta$Diagnosis)
Meta$Project <- as.factor(Meta$Project)

# Pathways
Path <- read.csv(paste0(D.dir, "pathways.csv"), row.names = 1, header = TRUE)

# Taxonomy
Taxo <- read.table(paste0(D.dir, "taxonomy.txt"), row.names = 1, header = TRUE)

# Round
TaxoR <- round(Taxo*100, 0)
```

# Phyloseq

We make object phyloseq

```{r}
# OTU table
OTU <- otu_table(as.matrix(TaxoR), taxa_are_rows = TRUE)

# SAM table
SAM <- sample_data(Meta)

# Phyloseq object
physeq <- merge_phyloseq(phyloseq(OTU), SAM)
```

# Results

## Abundance

Abundance top

```{r}
# Define the number top taxa
ntop <- 10

# Select the top taxa based on their abundance sums
top <- sort(taxa_sums(physeq), T)[1:ntop]

# Filter the phyloseq object to include only the top taxa
top <- prune_taxa(names(top), physeq)

# Convert to dataframe
topDF <- psmelt(top)

# Add a "species" column that is equal to OTU
topDF$Species <- topDF$OTU

# Plot 
topplot <- ggplot(data = topDF, aes(x = Diagnosis, y = Abundance, fill = Species)) +
    geom_bar(stat = "identity", position = "stack") + #position = "fill") + 
    # Scale
    labs(title = "Absolute abundance", x = "Diagnosis", 
         y = "Absolute abundance", fill = "Species") +
    # Theme
    theme_classic()
topplot

ggsave(filename = paste0(R.dir, "Abundance/", data, "-AbsAbun-Top10.svg"), 
       plot = topplot, width = 10, height = 10, units = "in")
```

## Diversity

Alpha

```{r}
# Alpha diversity measures
measures <- c("Observed", "Chao1", "Shannon", "Simpson")

# Plot
alpha <- plot_richness(physeq = physeq, measures = measures,
                       color = "Diagnosis", x = "Diagnosis", 
                       title = "Alpha Diversity Indices") +
  labs(x = "Diagnosis", y = "Alpha Diversity Measures") +
  theme_classic() +
  geom_boxplot(aes(fill = Diagnosis), alpha = 0.7) + 
  scale_color_brewer(palette = "Set1") + 
  scale_fill_brewer(palette = "Set1")
alpha

ggsave(filename = paste0(R.dir, "Diversity/", data, "-AlphaDiversity.svg"), 
       plot = alpha, width = 15, height = 10, units = "in")
```

Beta

```{r}
Ord <- ordinate(physeq = physeq, method = "PCoA", distance = "bray")
beta <- plot_ordination(physeq = physeq, ordination = Ord, color = "Diagnosis") +
    theme_classic() +
    geom_point(size = 3) +
    stat_ellipse(aes(group = Diagnosis, linetype = Diagnosis), type = "t", size = 0.75) +
    #geom_text(mapping = aes(label = Diagnosis), size = 4, vjust = 2, hjust = 1) + 
    geom_vline(xintercept = 0) +
    geom_hline(yintercept = 0) +
    labs(title = "Beta diversity (PCoA Bray-Curtis)") +
    scale_color_brewer(palette = "Set1") + 
    scale_fill_brewer(palette = "Set1")
beta
```

Escalamiento

```{r}
# Percentage
p.physeq <- transform_sample_counts(physeq = physeq, fun = function(x) x/sum(x))
p.Ord <- ordinate(physeq = p.physeq, method = "PCoA", distance = "bray")
plot_ordination(physeq = p.physeq, ordination = p.Ord, color = "Diagnosis") +
  stat_ellipse(aes(group = Diagnosis, linetype = Diagnosis), type = "t", size = 0.75) +
  theme_classic()
# Log
l.physeq <- transform_sample_counts(physeq = physeq, fun = function(x) log(x+1))
l.Ord <- ordinate(physeq = l.physeq, method = "PCoA", distance = "bray")
plot_ordination(physeq = l.physeq, ordination = l.Ord, color = "Diagnosis") +
  stat_ellipse(aes(group = Diagnosis, linetype = Diagnosis), type = "t", size = 0.75) +
  theme_classic()
# Square root
s.physeq <- transform_sample_counts(physeq = physeq, fun = function(x) sqrt(x))
s.Ord <- ordinate(physeq = s.physeq, method = "PCoA", distance = "bray")
plot_ordination(physeq = s.physeq, ordination = s.Ord, color = "Diagnosis")+
  stat_ellipse(aes(group = Diagnosis, linetype = Diagnosis), type = "t", size = 0.75) +
  theme_classic()
```

## Differential Taxa

### Edge R

```{r}
OTU <- physeq@otu_table@.Data
SAM <- as.data.frame(physeq@sam_data[,c("Diagnosis", "SampleID")])
SAM <- SAM[order(SAM$SampleID), ]
col.names <- colnames(OTU)
new.col.names <- as.character(SAM$Diagnosis[match(col.names, SAM$SampleID)])
colnames(OTU) <- ifelse(is.na(new.col.names), col.names, new.col.names)

counts <- as.data.frame(OTU)
Original <- dim(counts)
counts <- counts[rowSums(cpm(counts) >= 3) >= 1, ] 
Filtering <- dim(counts) 
print(paste("The percentage remaining after filtering is",
            round(Filtering[1]/Original[1]*100, 2))) 
```

Declare the variable where the DGEList object is stored

```{r}
grp <- new.col.names
```

We generate the DGEList object

```{r}
dge <- DGEList(counts = counts, group = grp)
```

Calculate a normalization factor.

```{r}
dge <- calcNormFactors(dge)
```

Estimate the dispersion

```{r}
dge <- estimateCommonDisp(dge)
```

GLMTagwise dispersion for multifactor experiments

````{r, results="hide"}
dge <- estimateGLMTagwiseDisp(dge)
```

We plot the dispersion of the total libraries, after the GMLTagwise analysis.

```{r}
plotBCV(dge)
# The lower the dispersion value the better since we expect them to be not 
# too sparse, i.e., the more clustered the better.
```

Matrix design

```{r}
design <- model.matrix(~ 0 + group, data = dge$samples) 
colnames(design) <- levels(dge$samples$group)
```

Comparisons by designing a vector of contrasts.

```{r}
fit <- glmQLFit(dge, design)
my.contrasts = makeContrasts(CDvsHE = CD - Healthy,
                             OBvsHE = Obese - Healthy,
                             UCvsHE = UC - Healthy,
                             levels = design)
```

DT Analysis

```{r}
# We declare the variables for each contrast
contrasts <- colnames(my.contrasts)

# In the for loop we will enter different values of FDRs
FDRs <- c(0.05, 0.005)

DTsE <- paste0(R.dir, "DifferentialTaxa/EdgeR/")
dir.create(DTsE, showWarnings = FALSE)

# Cycle for
for (contrast in contrasts){
  tryCatch({
    # Comparison
    # contrast <- "CDvsHE"
    # Genewise Negative Binomial
    qlf <- glmQLFTest(fit, contrast = my.contrasts[,contrast]) 
    
    # Table of the Top Differentially Expressed Taxones
    topTags(qlf)
    tabtop <- topTags(qlf, n = Inf)$table
    
    # DT
    for (fdr in FDRs){
      tryCatch({
        # fdr <- 0.05
        # Directories
        DTs.FDR <- paste0(DTsE, "FDR_", fdr, "/")
        dir.create(DTs.FDR, showWarnings = FALSE)
        DTs.FDR.T <- paste0(DTs.FDR, "Tablas/")
        dir.create(DTs.FDR.T, showWarnings = FALSE)
        DTs.FDR.P <- paste0(DTs.FDR, "Plots/")
        dir.create(DTs.FDR.P, showWarnings = FALSE)
        
        # DT.Taxones
        taxones <- rownames(tabtop)[tabtop$FDR <= fdr]
        length(taxones)
        
        # Save table DTs
        name <- paste0("DT-", contrast, "-FDR_", fdr) 
        write.table(tabtop[taxones,], file = paste0(DTs.FDR.T, name, ".txt"), 
                    row.names = TRUE, col.names = NA, quote = FALSE, sep = "\t")
        
        # Plots DTs
        plotV <- EnhancedVolcano(tabtop[tabtop$FDR <= fdr,], colAlpha = 1,
                                 title = paste0("Volcano plot of ", name),
                                 lab = taxones, x = "logFC", y = "PValue",
                                 gridlines.major = FALSE, gridlines.minor = FALSE)
        
        # Save plot
        ggsave(filename = paste0(DTs.FDR.P, name, "_Volcano.svg"), plot = plotV, 
               device = "svg", width = 15, height = 15, units = "in")
        
        # Separate up and down taxones
        updown <- read.table(paste0(DTs.FDR.T, name, ".txt"), header = TRUE, 
                        row.names = 1, sep = "\t", comment.char = "")
        write.table(updown[updown$logFC <= 0,], 
                    file = paste0(DTs.FDR.T, "down.", name, ".txt"), 
                    row.names = TRUE, col.names = NA, quote = FALSE, sep = "\t")
        write.table(updown[updown$logFC >= 0,], 
                    file = paste0(DTs.FDR.T, "up.", name, ".txt"), 
                    row.names = TRUE, col.names = NA, quote = FALSE, sep = "\t")
        
        # Applying a logFC > 1
        # Directories
        DTs.FDR.Log <- paste0(DTsE, "FDR_", fdr, "_logFC/")
        dir.create(DTs.FDR.Log, showWarnings = FALSE)
        DTs.FDR.Log.T <- paste0(DTs.FDR.Log, "Tablas/")
        dir.create(DTs.FDR.Log.T, showWarnings = FALSE)
        DTs.FDR.Log.P <- paste0(DTs.FDR.Log, "Plots/")
        dir.create(DTs.FDR.Log.P, showWarnings = FALSE)
        
        # DT.Taxones
        taxon<- rownames(tabtop)[tabtop$FDR <= fdr & abs(tabtop$logFC) >= 1] # de Taxones
        length(taxon)
        
        # Save table DTs
        name <- paste0("DT-", contrast, "-FDR_", fdr, "_logFC") # DTs
        write.table(tabtop[taxon,], file = paste0(DTs.FDR.Log.T, name, ".txt"), 
                    row.names = TRUE, col.names = NA, quote = FALSE, sep = "\t")
        
        # Plots DTs
        plotV <- EnhancedVolcano(tabtop[tabtop$FDR <= fdr & abs(tabtop$logFC) >= 1,], 
                                 title = paste0("Volcano plot of ", name),
                                 lab = taxon, x = "logFC", y = "PValue", colAlpha = 1,
                                 gridlines.major = FALSE, gridlines.minor = FALSE)
        
        # Save plot
        ggsave(filename = paste0(DTs.FDR.Log.P, name, "_Volcano.svg"), plot = plotV, 
               device = "svg", width = 15, height = 15, units = "in")

        # Separate up and down taxones
        updown <- read.table(paste0(DTs.FDR.Log.T, name, ".txt"), 
                             header = TRUE, row.names = 1, sep = "\t", comment.char="")
        write.table(updown[updown$logFC <= 0,], 
                    file = paste0(DTs.FDR.Log.T, "down.", name, ".txt"), 
                    row.names = TRUE, col.names = NA, quote = FALSE, sep = "\t")
        write.table(updown[updown$logFC >= 0,], 
                    file = paste0(DTs.FDR.Log.T, "up.", name, ".txt"), 
                    row.names = TRUE, col.names = NA, quote = FALSE, sep = "\t")
      }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
    }
  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}
```

DT Matrix for each condition

```{r}
conditions <- c("FDR_0.05", "FDR_0.05_logFC",
                "FDR_0.005", "FDR_0.005_logFC")

for (fdr in conditions){
  list <- list()
  id <- c()
  # fdr <- "FDR_0.05"
  
  # Tables of FDRs only
  DTs.FDR <- paste0(DTsE, fdr, "/")
  DTs.FDR.T <- paste0(DTs.FDR, "Tablas/")
  
  # Each of the filenames with the FDR_ pattern is stored
  countFiles <- list.files(path = DTs.FDR.T, 
                           pattern = paste0(fdr, ".txt"))
  
  for (count in countFiles) {
    # count <- "DT-IPT7vsWT7-FDR_0.050.txt"
    tab <- read.table(file = paste0(DTs.FDR.T, count), 
                      header = TRUE, sep="\t", quote= "", 
                      row.names=1, comment.char="")
    list[[count]] <- tab
    id <- unique(c(id, rownames(tab)))
  }
  
  for (type in c("logFC", "logCPM", "PValue", "FDR")){
    # type <- "logFC"
    table <- data.frame(row.names = id)
    for (treat in names(list)) {
      temptable <- list[[treat]]
      table[rownames(temptable), treat] <- temptable[, type]
      # print(paste("The sample", treat, "was successfully added"))
    }
    
    table[is.na(table)] = 0
    colnames(table) <- gsub(".txt", "", colnames(table))
    
    write.table(table, file = paste0(DTs.FDR, "Table-", type, "_DTs-", fdr, ".txt"), 
                row.names = TRUE, col.names = NA, quote = FALSE, sep = "\t")
  }
}
```

## Venn Diagram for DTs

We will perform the analysis for all conditions

```{r, warning = FALSE}
# No create file .log
futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")

# directory
VennE <- paste0(R.dir, "VennDiagramTaxa/")

# Combination contrasts
comb.contrats <- t(combn(contrasts, 2))

for (condition in conditions){
    # condition <- "FDR_0.05"
    # Directories
    Venn.FDR <- paste0(VennE, condition,"/") 
    dir.create(Venn.FDR, showWarnings = FALSE)
    VennPlot <- paste0(Venn.FDR, "Plots/") 
    dir.create(VennPlot, showWarnings = FALSE)
    VennList <- paste0(Venn.FDR, "Lista/") 
    dir.create(VennList, showWarnings = FALSE)
    
    # Each file name is stored with the pattern "FDR_0.050.txt"
    countFiles <- list.files(path = paste0(DTsE, condition, "/Tablas/"),
                             pattern = paste0(condition, ".txt"))
    
    # We make a list of taxones of each contrast in the given condition
    list_Venn <- list()
    for (count in countFiles){
        # count <- "DT-OBvsHe-FDR_0.05.txt"
        name <- sub(paste0("-", condition, ".txt"), "", count)
        data <- read.table(paste0(DTsE, condition, "/Tablas/", count), 
                           header = TRUE, sep = "\t", quote = "", 
                           row.names = 1, comment.char = "")
        list_Venn[[name]] <- data
    }
    
    # We analyze the various possible combinations of 2 element
    for (i in 1:nrow(comb.contrats)){
        # i <- 3
        repA <- comb.contrats[i,][1]
        repB <- comb.contrats[i,][2]
        
        for (ud in c("", "up.", "down.")){
            # ud <- ""
            vennlist <- list(repA = rownames(list_Venn[[paste0(ud, "DT-", repA)]]),
                             repB = rownames(list_Venn[[paste0(ud, "DT-", repB)]]))
            names(vennlist) <- c(paste0(ud, repA), paste0(ud, repB))
            
            if (length(vennlist[[1]]) != 0 && length(vennlist[[2]]) != 0){
                # Venn object
                ventemp <- venn(vennlist, show.plot = FALSE)
                ventemp <- attributes(ventemp)$intersections
                ventemp <- t(do.call(rbind, ventemp))
                
                # Verifies whether or not intersection exists
                if (sum(table(unique(ventemp[,3]))) > 0 && dim(ventemp)[2] == 3){
                    tryCatch({
                        # Graphs
                        venn.diagram(x = vennlist, 
                                     # Names
                                     category.names = names(vennlist), 
                                     main = "Venn Diagram", main.cex = 20,
                                     main.fontfamily = "sans",
                                     sub = paste0(ud, repA, " vs ", ud, repB), 
                                     sub.cex = 15, sub.fontfamily = "sans",
                                     # Output
                                     filename = paste0(VennPlot, ud, "DT_", 
                                                       repA, "-vs-", repB, ".svg"), 
                                     output = FALSE, 
                                     # Output features
                                     imagetype = "svg", height = 100, 
                                     width = 100, resolution = 300,
                                     # Circles
                                     lwd = 15, 
                                     col = c("#440154ff", "#21908dff"),
                                     fill = c(alpha("#440154ff",0.3),
                                              alpha("#21908dff",0.3)),
                                     # Numbers
                                     cex = 15, fontfamily = "sans",
                                     # Set names
                                     cat.pos = c(315, 45), 
                                     cat.dist = c(0.05, 0.05), 
                                     cat.cex = 15, 
                                     cat.default.pos = "outer", 
                                     cat.fontfamily = "sans")
                      
                        # Save table with all
                        write.table(ventemp, file = paste0(VennList, "ListVenn-", 
                                                           ud, "DT_", repA, "-vs-", 
                                                           repB, ".txt"),
                                    row.names = FALSE, quote = FALSE, 
                                    sep = "\t", col.names = TRUE)
                        
                        # Cycle to save tables
                        for (col in 1:length(colnames(ventemp))) {
                            # Crea el dataframe
                            df <- data.frame(unique(ventemp[, col]))
                            names(df) <- colnames(ventemp)[col]
                            names(df) <- gsub(":", "&", names(df))
                            
                            # Guarda el dataframe
                            write.table(df, file = paste0(VennList, "ListVenn-", 
                                                          ud, "DT_", repA, "-vs-", 
                                                          repB, "_only_", 
                                                          names(df), ".txt"),
                                        row.names = FALSE, quote = FALSE, 
                                        sep = "\t", col.names = TRUE)
                        }
                    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
                }
            }
        }
    }
    
    # Venn of 3 contrast join
    for (ud in c("", "up.", "down.")){
        # ud <- "up."
        x <- contrasts[1]
        y <- contrasts[2]
        z <- contrasts[3]
        
        vennlist <- list(x = rownames(list_Venn[[paste0(ud, "DT-", x)]]),
                         y = rownames(list_Venn[[paste0(ud, "DT-", y)]]),
                         z = rownames(list_Venn[[paste0(ud, "DT-", z)]]))
        
        names(vennlist) <- c(paste0(ud, x), paste0(ud, y), paste0(ud, z))
        
        if (length(vennlist[[1]]) != 0 && length(vennlist[[2]]) != 0 && length(vennlist[[3]]) != 0 ){
            ventemp <- venn(vennlist, show.plot = FALSE)
            ventemp <- attributes(ventemp)$intersections
            ventemp <- t(do.call(rbind, ventemp))
            
            if (sum(table(unique(ventemp[,7]))) > 0 && dim(ventemp)[2] == 7){
                tryCatch({
                    venn.diagram(x = vennlist, 
                                 # Names
                                 category.names = names(vennlist), 
                                 main = "Venn Diagram", main.cex = 20, 
                                 sub = paste0(ud, x, " vs ", ud, y, " vs ", ud, z), 
                                 sub.cex = 15, main.fontfamily = "sans", 
                                 sub.fontfamily = "sans",
                                 # Output
                                 filename = paste0(VennPlot, ud, "DT_", x, 
                                                   "-vs-", y, "-vs-", z, ".svg"), 
                                 output = FALSE, 
                                 # Output features
                                 imagetype = "svg", height = 100, 
                                 width = 100, resolution = 300,
                                 # Circles
                                 lwd = 15, 
                                 col = c("#440154ff", "#21908dff", "#fde725ff"),
                                 fill = c(alpha("#440154ff",0.3), 
                                          alpha("#21908dff",0.3), 
                                          alpha("#fde725ff",0.3)),
                                 # Numbers
                                 cex = 15, fontfamily = "sans",
                                 # Set names
                                 cat.pos = c(-27, 27, 135), 
                                 cat.dist = c(0.05, 0.05, 0.05), 
                                 cat.cex = 15, 
                                 cat.default.pos = "outer", 
                                 cat.fontfamily = "sans")
                  
                    # Save table with all
                    write.table(ventemp, file = paste0(VennList, "ListVenn-",
                                                       ud, "DT_", x, "-vs-", y, 
                                                       "-vs-", z, ".txt"),
                                row.names = FALSE, quote = FALSE, 
                                sep = "\t", col.names = TRUE)
                    
                    # Cycle to save tables
                    for (col in 1:length(colnames(ventemp))) {
                        # Crea el dataframe
                        df <- data.frame(unique(ventemp[, col]))
                        names(df) <- colnames(ventemp)[col]
                        names(df) <- gsub(":", "&", names(df))
                        # Guarda el dataframe
                        write.table(df, file = paste0(VennList, "ListVenn-", 
                                                      ud, "DT_", x, "-vs-", 
                                                      y, "-vs-", z, "_only_", 
                                                      names(df), ".txt"),
                                    row.names = FALSE, quote = FALSE, 
                                    sep = "\t", col.names = TRUE)
                    }
                }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
            }
        }
    }
}
```

## Differential Function Pathways

```{r}
pathway <- Path*1000000000
col.names <- colnames(pathway)
new.col.names <- as.character(SAM$Diagnosis[match(col.names, SAM$SampleID)])
colnames(pathway) <- ifelse(is.na(new.col.names), col.names, new.col.names)

countsF <- pathway
Original <- dim(countsF) 
countsF <- countsF[rowSums(cpm(countsF) >= 1) >= 1, ] 
Filtering <- dim(countsF) 
print(paste("The percentage remaining after filtering is",
            round(Filtering[1]/Original[1]*100, 2))) 
```

Declare the variable where the DGEList object is stored

```{r}
grpF <- new.col.names
```

DGEList object

```{r}
dgeF <- DGEList(counts = countsF, group = grpF)
```

Calculate a normalization factor.

```{r}
dgeF <- calcNormFactors(dgeF)
```

Estimate the dispersion

```{r}
dgeF <- estimateCommonDisp(dgeF)
```

GLMTagwise dispersion for multifactor experiments

````{r, results="hide"}
dgeF <- estimateGLMTagwiseDisp(dgeF)
```

Matrix design

```{r}
designF <- model.matrix(~ 0 + group, data = dgeF$samples) 
colnames(designF) <- levels(dgeF$samples$group)
```

Make comparisons 

```{r}
fitF <- glmQLFit(dgeF, designF)
# Subtraction is always "Treatment - Control".
my.contrasts = makeContrasts(CDvsHE = CD - Healthy,
                             OBvsHE = Obese - Healthy,
                             UCvsHE = UC - Healthy,
                             levels = designF)
```

Differential Function Analysis

```{r}
# We declare the variables for each contrast
contrasts <- colnames(my.contrasts)

# In the for loop we will enter different values of FDRs
FDRs <- c(0.05, 0.005)

DFun.P <- paste0(R.dir, "DifferentialFunction/")

# Cycle for
for (contrast in contrasts){
  tryCatch({
    # Comparison
    # contrast <- "CDvsHE"
    
    # Genewise Negative Binomial
    qlf <- glmQLFTest(fitF, contrast = my.contrasts[,contrast]) 
    
    # Table of the Top Differentially Expressed Taxones
    topTags(qlf)
    tabtop <- topTags(qlf, n = Inf)$table
    
    # DF
    for (fdr in FDRs){
      tryCatch({
        # fdr <- 0.05
        # Directories
        DFun.FDR <- paste0(DFun.P, "FDR_", fdr, "/")
        dir.create(DFun.FDR, showWarnings = FALSE)
        DFun.FDR.T <- paste0(DFun.FDR, "Tablas/")
        dir.create(DFun.FDR.T, showWarnings = FALSE)
        DFun.FDR.P <- paste0(DFun.FDR, "Plots/")
        dir.create(DFun.FDR.P, showWarnings = FALSE)
        
        # DF.Taxones
        taxones <- rownames(tabtop)[tabtop$FDR <= fdr]
        length(taxones)
        df.functions <- tabtop[taxones,] %>% drop_na()
        
        # Save table DFun
        name <- paste0("DF-", contrast, "-FDR_", fdr) 
        write.table(df.functions, file = paste0(DFun.FDR.T, name, ".txt"), 
                    row.names = TRUE, col.names = NA, quote = FALSE, sep = "\t")
        
        # Plots DFun
        plotV <- EnhancedVolcano(df.functions, colAlpha = 1,
                                 title = paste0("Volcano plot of ", name),
                                 lab = taxones, x = "logFC", y = "PValue",
                                 gridlines.major = FALSE, gridlines.minor = FALSE)
        
        # Save plot
        ggsave(filename = paste0(DFun.FDR.P, name, "_Volcano.svg"), plot = plotV, 
               device = "svg", width = 15, height = 15, units = "in")
        
        # Separate up and down taxones
        df.fun.down <- df.functions[df.functions$logFC <= 0,] %>% drop_na()
        write.table(df.fun.down, file = paste0(DFun.FDR.T, "down.", name, ".txt"), 
                    row.names = TRUE, col.names = NA, quote = FALSE, sep = "\t")
        df.fun.up <- df.functions[df.functions$logFC >= 0,] %>% drop_na()
        write.table(df.fun.up, file = paste0(DFun.FDR.T, "up.", name, ".txt"), 
                    row.names = TRUE, col.names = NA, quote = FALSE, sep = "\t")
        
        # Applying a logFC > 1
        # Directories
        DFun.FDR.Log <- paste0(DFun.P, "FDR_", fdr, "_logFC/")
        dir.create(DFun.FDR.Log, showWarnings = FALSE)
        DFun.FDR.Log.T <- paste0(DFun.FDR.Log, "Tablas/")
        dir.create(DFun.FDR.Log.T, showWarnings = FALSE)
        DFun.FDR.Log.P <- paste0(DFun.FDR.Log, "Plots/")
        dir.create(DFun.FDR.Log.P, showWarnings = FALSE)
        
        # DF.Taxones
        taxon <- rownames(tabtop)[tabtop$FDR <= fdr & abs(tabtop$logFC) >= 1] # de Taxones
        length(taxon)
        df.functions <- tabtop[taxon,] %>% drop_na()
        
        # Save table DFun
        name <- paste0("DF-", contrast, "-FDR_", fdr, "_logFC") # DFun
        write.table(df.functions, file = paste0(DFun.FDR.Log.T, name, ".txt"), 
                    row.names = TRUE, col.names = NA, quote = FALSE, sep = "\t")
        
        # Plots DFun
        plotV <- EnhancedVolcano(df.functions, 
                        lab = taxon, x = "logFC", y = "PValue", colAlpha = 1,
                        title = paste0("Volcano plot of ", name),
                        gridlines.major = FALSE, gridlines.minor = FALSE)
        
        # Save plot
        ggsave(filename = paste0(DFun.FDR.Log.P, name, "_Volcano.svg"), plot = plotV, 
               device = "svg", width = 15, height = 15, units = "in")
        
        # Separate up and down taxones
        df.fun.down <- df.functions[df.functions$logFC <= 0,] %>% drop_na()
        write.table(df.fun.down, file = paste0(DFun.FDR.Log.T, "down.", name, ".txt"), 
                    row.names = TRUE, col.names = NA, quote = FALSE, sep = "\t")
        df.fun.up <- df.functions[df.functions$logFC >= 0,] %>% drop_na()
        write.table(df.fun.up, file = paste0(DFun.FDR.Log.T, "up.", name, ".txt"), 
                    row.names = TRUE, col.names = NA, quote = FALSE, sep = "\t")
      }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
    }
  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}
```

DF Matrix for each condition

```{r}
conditions <- c("FDR_0.05", "FDR_0.05_logFC",
                "FDR_0.005", "FDR_0.005_logFC")

for (fdr in conditions){
  list <- list()
  id <- c()
  # fdr <- "FDR_0.05"
  
  # Tables of FDRs only
  DFun.FDR <- paste0(DFun.P, fdr, "/")
  DFun.FDR.T <- paste0(DFun.FDR, "Tablas/")
  
  # Each of the filenames with the FDR_ pattern is stored
  countFiles <- list.files(path = DFun.FDR.T, 
                           pattern = paste0(fdr, ".txt"))
  
  for (count in countFiles) {
    # count <- "DF-IPT7vsWT7-FDR_0.050.txt"
    tab <- read.table(file = paste0(DFun.FDR.T, count), 
                      header = TRUE, sep="\t", quote= "", 
                      row.names=1, comment.char="")
    list[[count]] <- tab
    id <- unique(c(id, rownames(tab)))
  }
  
  for (type in c("logFC", "logCPM", "PValue", "FDR")){
    # type <- "logFC"
    table <- data.frame(row.names = id)
    for (treat in names(list)) {
      temptable <- list[[treat]]
      table[rownames(temptable), treat] <- temptable[, type]
      # print(paste("The sample", treat, "was successfully added"))
    }
    
    table[is.na(table)] = 0
    colnames(table) <- gsub(".txt", "", colnames(table))
    write.table(table, file = paste0(DFun.FDR, "Table-", type, "_DFun-", fdr, ".txt"), 
                row.names = TRUE, col.names = NA, quote = FALSE, sep = "\t")
  }
}
```  

## Venn Diagram Functions

```{r, warning = FALSE}
# No create file .log
futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")

# Directory
VennE <- paste0(R.dir, "VennDiagramFunction/")

# Combination contrasts
comb.contrats <- t(combn(contrasts, 2))

for (condition in conditions){
    # condition <- "FDR_0.05"
    # Directories
    Venn.FDR <- paste0(VennE, condition,"/") 
    dir.create(Venn.FDR, showWarnings = FALSE)
    VennPlot <- paste0(Venn.FDR, "Plots/") 
    dir.create(VennPlot, showWarnings = FALSE)
    VennList <- paste0(Venn.FDR, "Lista/") 
    dir.create(VennList, showWarnings = FALSE)
    
    # Each file name is stored with the pattern "FDR_0.050.txt"
    countFiles <- list.files(path = paste0(DFun.P, condition, "/Tablas/"),
                             pattern = paste0(condition, ".txt"))
    
    # We make a list of taxones of each contrast in the given condition
    list_Venn <- list()
    for (count in countFiles){
        # count <- "DF-OBvsHe-FDR_0.05.txt"
        name <- sub(paste0("-", condition, ".txt"), "", count)
        data <- read.table(paste0(DFun.P, condition, "/Tablas/", count), 
                           header = TRUE, sep = "\t", quote = "", 
                           row.names = 1, comment.char = "")
        list_Venn[[name]] <- data
    }
    
    # We analyze the various possible combinations of 2 element
    for (i in 1:nrow(comb.contrats)){
        # i <- 3
        repA <- comb.contrats[i,][1]
        repB <- comb.contrats[i,][2]
        
        for (ud in c("", "up.", "down.")){
            # ud <- ""
            vennlist <- list(repA = rownames(list_Venn[[paste0(ud, "DF-", repA)]]),
                             repB = rownames(list_Venn[[paste0(ud, "DF-", repB)]]))
            
            names(vennlist) <- c(paste0(ud, repA), paste0(ud, repB))
            
            if (length(vennlist[[1]]) != 0 && length(vennlist[[2]]) != 0){
                # Venn object
                ventemp <- venn(vennlist, show.plot = FALSE)
                ventemp <- attributes(ventemp)$intersections
                ventemp <- t(do.call(rbind, ventemp))
                
                # Verifies whether or not intersection exists
                if (sum(table(unique(ventemp[,3]))) > 0 && dim(ventemp)[2] == 3){
                    tryCatch({
                        # Graphs
                        venn.diagram(x = vennlist, 
                                     # Names
                                     category.names = names(vennlist), 
                                     main = "Venn Diagram", main.cex = 20,
                                     main.fontfamily = "sans",
                                     sub = paste0(ud, repA, " vs ", ud, repB), 
                                     sub.cex = 15, sub.fontfamily = "sans",
                                     # Output
                                     filename = paste0(VennPlot, ud, "DF_", 
                                                       repA, "-vs-", repB, ".svg"), 
                                     output = FALSE, 
                                     # Output features
                                     imagetype = "svg", height = 100, 
                                     width = 100, resolution = 300,
                                     # Circles
                                     lwd = 15, 
                                     col = c("#440154ff", "#21908dff"),
                                     fill = c(alpha("#440154ff",0.3),
                                              alpha("#21908dff",0.3)),
                                     # Numbers
                                     cex = 15, fontfamily = "sans",
                                     # Set names
                                     cat.pos = c(315, 45), 
                                     cat.dist = c(0.05, 0.05), 
                                     cat.cex = 15, 
                                     cat.default.pos = "outer", 
                                     cat.fontfamily = "sans")
                      
                        # Save table with all
                        write.table(ventemp, file = paste0(VennList, "ListVenn-", 
                                                           ud, "DF_", repA, "-vs-", 
                                                           repB, ".txt"),
                                    row.names = FALSE, quote = FALSE, 
                                    sep = "\t", col.names = TRUE)
                        
                        # Cycle to save tables
                        for (col in 1:length(colnames(ventemp))) {
                            # Crea el dataframe
                            # col <- 1
                            df <- data.frame(unique(ventemp[, col]))
                            names(df) <- colnames(ventemp)[col]
                            names(df) <- gsub(":", "&", names(df))
                            # Guarda el dataframe
                            write.table(df, file = paste0(VennList, "ListVenn-", 
                                                          ud, "DF_", repA, "-vs-", 
                                                          repB, "_only_", 
                                                          names(df), ".txt"),
                                        row.names = FALSE, quote = FALSE, 
                                        sep = "\t", col.names = TRUE)
                        }
                    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
                }
            }
        }
    }
    
    # Venn of 3 contrast join
    for (ud in c("", "up.", "down.")){
        # ud <- "up."
        x <- contrasts[1]
        y <- contrasts[2]
        z <- contrasts[3]
        
        vennlist <- list(x = rownames(list_Venn[[paste0(ud, "DF-", x)]]),
                         y = rownames(list_Venn[[paste0(ud, "DF-", y)]]),
                         z = rownames(list_Venn[[paste0(ud, "DF-", z)]]))
        
        names(vennlist) <- c(paste0(ud, x), paste0(ud, y), paste0(ud, z))
        
        if (length(vennlist[[1]]) != 0 && length(vennlist[[2]]) != 0 && length(vennlist[[3]]) != 0 ){
            ventemp <- venn(vennlist, show.plot = FALSE)
            ventemp <- attributes(ventemp)$intersections
            ventemp <- t(do.call(rbind, ventemp))
            
            if (sum(table(unique(ventemp[,7]))) > 0 && dim(ventemp)[2] == 7){
                tryCatch({
                    venn.diagram(x = vennlist, 
                                 # Names
                                 category.names = names(vennlist), 
                                 main = "Venn Diagram", main.cex = 20, 
                                 sub = paste0(ud, x, " vs ", ud, y, " vs ", ud, z), 
                                 sub.cex = 15, main.fontfamily = "sans", 
                                 sub.fontfamily = "sans",
                                 # Output
                                 filename = paste0(VennPlot, ud, "DF_", x, 
                                                   "-vs-", y, "-vs-", z, ".svg"), 
                                 output = FALSE, 
                                 # Output features
                                 imagetype = "svg", height = 100, 
                                 width = 100, resolution = 300,
                                 # Circles
                                 lwd = 15, 
                                 col = c("#440154ff", "#21908dff", "#fde725ff"),
                                 fill = c(alpha("#440154ff",0.3), 
                                          alpha("#21908dff",0.3), 
                                          alpha("#fde725ff",0.3)),
                                 # Numbers
                                 cex = 15, fontfamily = "sans",
                                 # Set names
                                 cat.pos = c(-27, 27, 135), 
                                 cat.dist = c(0.05, 0.05, 0.05), 
                                 cat.cex = 15, 
                                 cat.default.pos = "outer", 
                                 cat.fontfamily = "sans")
                  
                    # Save table with all
                    write.table(ventemp, file = paste0(VennList, "ListVenn-",
                                                       ud, "DF_", x, "-vs-", y, 
                                                       "-vs-", z, ".txt"),
                                row.names = FALSE, quote = FALSE, 
                                sep = "\t", col.names = TRUE)
                    
                    # Cycle to save tables
                    for (col in 1:length(colnames(ventemp))) {
                        # Crea el dataframe
                        df <- data.frame(unique(ventemp[, col]))
                        names(df) <- colnames(ventemp)[col]
                        names(df) <- gsub(":", "&", names(df))
                        # Guarda el dataframe
                        write.table(df, file = paste0(VennList, "ListVenn-", 
                                                      ud, "DF_", x, "-vs-", 
                                                      y, "-vs-", z, "_only_", 
                                                      names(df), ".txt"),
                                    row.names = FALSE, quote = FALSE, 
                                    sep = "\t", col.names = TRUE)
                    }
                }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
            }
        }
    }
}
```